% ******************************************************
% Acknowledgments
% ******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
        They didn't have much trouble \\
        teaching the ape to write poems: \\
        first they strapped him into a chair, \\
        then tied the pencil around his hand \\
        (the paper had already been nailed down). \\
        Then Dr.\ Bluespire leaned over his shoulder \\
        and whispered into his ear: \\
        You look like a god sitting there. \\
        Why don't you try writing something?} \\ \medskip
    --― James Tate
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgments}

Many thanks are in place for the successful completion of this thesis.
First I would like to thank my academic advisor Joel Wallman for the guidance during my bumpy career as a graduate student.
In addition, thank you to the following professors to helping me complete my studies: John Watrous, Achim Kempf, Michael Waite, Brian Ingalls, and Michael Brannan.
Whether it was sharing details about your personal career, asking probing questions, or offering time and having supportive conversation despite not having to: thank you.

Thank you to Joel's research group for helping me deal with Joel's departure: Darian Mclaren, Anthony Chytros, Matthew Graydon, Stefanie Beale, Sam Ferracin, and Joshua Skanes-Norman.
I would also like to thank my many classmates without which remote classes would have been far less interesting and rewarding: Wilson Wu, Chelsea Komlo, Mohammad Ayyash, Nicholas Zutt, and Xiaoran Li.
A big thank you is also in order for Overleaf and in particular John Lees-Miller and Ryan Looney for allowing me to work part time and being extremely flexible with my hours.
It was great to continue working with the team\ldots and to supplement the measly graduate student salary.

Thank you Mom and Dad for letting me live in your house while we endured the brunt of the pandemic.
Thank you Diane for always having my back and being supportive throughout my graduate studies.
Thank you to my friends who were always open to discuss my struggles and triumphs: Kevin (both of you), Rafael, Ana, and Aimee.

\endgroup
