\documentclass[12pt]{article}

\setlength{\headheight}{14.5pt}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.2in]{geometry}
\usepackage{microtype}
\usepackage{fancyhdr}
\usepackage{palatino}
 
\pagestyle{fancy}
\fancyhf{}
\lhead{Cover Letter}
\rhead{Researcher - Internship}

\renewcommand{\baselinestretch}{1.42}
\setlength\parskip{0.5em}

\begin{document}
\noindent
To whom it may concern,

I am a second year PhD student at the Institute for Quantum Computing (IQC) who is interested in making quantum technologies a reality.
I began my work at IQC with Professor Joel Wallman working on quantum error correcting techniques and applications of Lie theory to quantum information.
At 1QBit I'd like to apply my quantum information knowledge, combined with my programming skills to work on the real problems we face in constructing quantum computers.

Having spent 4 years as a software developer at Overleaf, I developed strong programming abilities and learned how to be a core member of a fast moving team.
From finding and diagnosing bugs, to doing user research and shaping new features, I've run the gamut of the responsibilities a software developer may have.
My programming experience has lent me a new lens through which I view problems, and will allow me to work on software projects with less startup time

While at NYU I spent two summers in the Wright Lab at Yale where I got my hands dirty running experiments testing all components of our detector to optimize for our goal of searching for a new type of neutrino.
I also spent a summer in the mathematics department at NYU where I numerically computed solutions to partial differential equations modelling electromagnetic pulses travelling through a medium.
These summers gave me the understanding of how progress in science is made, and allowed me to experience both theoretical and experimental sciences.

With the advances transpiring in quantum technology I believe now to be an ideal time to be involved in creating a technology that will bring us into a new computing era.
The many serious political, social, and environmental effects of quantum technologies require it's creators to be both well versed, and well rounded quantum scientists.
I believe this opportunity at 1QBit to be an ideal place to expose me to the real problems we face in building quantum computers and will help me become the well rounded researcher I aspire to be.

\noindent
Best,

\parskip=0pt plus 1pt
Nathaniel T. Stemen
\end{document}